---
title: "Open Data Kit Quality Control Workflows"
subtitle: ""  
author: "Ernest Guevarra"
date: '16 November 2021'
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      titleSlideClass: ["right", "bottom"]
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)

library(zscorer)
library(nutricheckr)
library(dplyr)
library(knitr)
library(kableExtra)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_light(
  #base_color = "#509935",
  base_color = "#000000",
  base_font_size = "22px",
  title_slide_background_color = "#FFFFFF",
  title_slide_background_image = "images/ecohealth_title_background_4by3.png",
  title_slide_background_size = "contain",
  #title_slide_text_color = "#509935",
  title_slide_text_color = "#000000",
  background_image = "images/ecohealth_slide_background_4by3.png",
  header_font_google = google_font("Fira Sans"),
  text_font_google = google_font("Fira Sans Condensed"),
  code_font_google = google_font("Fira Mono"),
  code_font_size = "0.7rem",
  text_slide_number_font_size = "1em",
  link_color = "#509935"
)
```

### Outline

* Brief overview of **ODK Central** version 1.3 data review feature

* Demonstration of **ODK Central**'s data review feature

     - manual process
     - semi-automated process using R

* Points to remember when using **ODK Central**'s data review feature

---

### ODK Central v1.3 and data review feature

* Allows project manager/s to review submitted data from **ODK Central** and mark each row of data as either **approved**, **has issues**, or **rejected**.

* When marking a submission with any of the review labels mentioned, a comment can be written to describe the reasons/basis for the label assigned.

* Allows project manager/s to edit submitted data from **ODK Central** to resolve issues identified with the data submitted.

* If data has been edited, project manager/s can then re-review and re-assign a review label accordingly (i.e., from **rejected** to **approved**)

* creates history of edits done within **ODK Central** which can be used by project manager/s in the review process

---

### Demonstration of ODK Central data review feature - manual process

* login to your **ODK Central** EHA account - only accounts of those with project manager roles will have the review features available

* we will look at an example project on **anthropometric data collection for children 6-59 months old**

* this will be good for quick and simple **eyeball** checks of data

* more sophisticated and elaborate data checks will be too tedious and inefficient through this approach

---

### Demonstration of ODK Cenetral data review feature - semi-automated process using R

* access data in R using the `{ruODK}` package

* perform data quality review and checks (including more sophisticated/elaborate ones) on data using R functions for this purpose with the aim of identifying rows of data with issues

* login to your **ODK Central** EHA account, go to your project and then manually mark and comment on the rows of data identified with issues

---

### Access data in R using the `{ruODK}` package

```{r odksetup, eval = FALSE}
## Install ruODK from GitHub
remotes::install_github("ropensci/ruODK")

## Load ruODK
library(ruODK)

## Setup connection with ODK Central
ru_setup(
  svc = "https://odk.eha.io/v1/projects/1/forms/anthropometry.svc",
  un = "YOUR_USERNAME_HERE",
  pw = "YOUR_PASSWORD_HERE",
  tz = "GMT",
  odkc_version = "1.3"
)

## Retrieve test data
anthro <- odata_submission_get()

## View data
anthro
```

---

### Access data in R using the `{ruODK}` package

```{r odksetup_run, echo = FALSE, eval = TRUE}
library(ruODK)

ru_setup(
  svc = "https://odk.eha.io/v1/projects/1/forms/anthropometry.svc",
  un = Sys.getenv("ODK_USERNAME"),
  pw = Sys.getenv("ODK_PASSWORD"),
  tz = "GMT",
  odkc_version = "1.3",
  verbose = FALSE
)

anthro <- odata_submission_get()

anthro
```

---

background-image: url()
background-color: #FFFFFF

### Perform data quality review/checks

```{r data_review_1, echo = TRUE, eval = FALSE}
## Load anthropometric z-score calculator library and data checking library
library(zscorer)
library(nutricheckr)
library(dplyr)

## Calculate anthropometric z-scores
anthro_zscores <- anthro %>%
  mutate(age_days = anthropometry_age * (365.25 / 12)) %>%
  addWGSR( 
    sex = "anthropometry_sex", 
    firstPart = "anthropometry_weight",
    secondPart = "age_days",
    index = "wfa"
  ) %>%
  addWGSR(
    sex = "anthropometry_sex",
    firstPart = "anthropometry_height",
    secondPart = "age_days",
    index = "hfa"
  ) %>%
  addWGSR(
    sex = "anthropometry_sex",
    firstPart = "anthropometry_weight",
    secondPart = "anthropometry_height",
    index = "wfh"
  )

## View output
anthro_zscores
```

---

### Perform data quality review/checks

```{r data_review_1_run, echo = FALSE, eval = TRUE}
## Calculate anthropometric z-scores
anthro_zscores <- anthro %>%
  mutate(age_days = anthropometry_age * (365.25 / 12)) %>%
  addWGSR( 
    sex = "anthropometry_sex", 
    firstPart = "anthropometry_weight",
    secondPart = "age_days",
    index = "wfa"
  ) %>%
  addWGSR(
    sex = "anthropometry_sex",
    firstPart = "anthropometry_height",
    secondPart = "age_days",
    index = "hfa"
  ) %>%
  addWGSR(
    sex = "anthropometry_sex",
    firstPart = "anthropometry_weight",
    secondPart = "anthropometry_height",
    index = "wfh"
  )

anthro_zscores %>%
  select(
    -id, -start, -end, -today, -meta_instance_id, -meta_instance_name, 
    -system_submission_date, -system_submitter_id, -system_submitter_name, 
    -system_attachments_present, -system_attachments_expected, 
    -system_device_id, -system_edits, -odata_context
  ) %>%
  rename_with(.fn = ~gsub(pattern = "anthropometry_|admin_", replacement = "", .x))
```

---

### Perform data quality review/checks

```{r data_review_2, echo = TRUE, eval = FALSE}
## Flag z-scores
anthro_flags <- anthro_zscores %>%
  flag_who(hlaz = "hfaz", waz = "wfaz", whlz = "wfhz")

anthro_flags
```

```{r data_review_2_run, echo = FALSE, eval = TRUE}
## Flag z-scores
anthro_flags <- anthro_zscores %>%
  flag_who(hlaz = "hfaz", waz = "wfaz", whlz = "wfhz") %>%
    select(
    -id, -start, -end, -today, -meta_instance_id, -meta_instance_name, 
    -system_submission_date, -system_submitter_id, -system_submitter_name, 
    -system_attachments_present, -system_attachments_expected, 
    -system_device_id, -system_edits, -odata_context
  ) %>%
  rename_with(.fn = ~gsub(pattern = "anthropometry_|admin_", replacement = "", .x))

tibble::tibble(anthro_flags)
```

---

### Perform data quality review/checks

```{r data_review_3, echo = TRUE, eval = FALSE}
## Subset anthropometric data to flagged rows
anthro_for_checking <- anthro_flags %>%
  filter(flag != 0)

anthro_for_checking
```

```{r data_review_3_run, echo = FALSE, eval = TRUE}
## Subset anthropometric data to flagged rows
anthro_for_checking <- anthro_flags %>%
  filter(flag != 0)

tibble::tibble(anthro_for_checking)
```

---

class: center, middle

# Questions?

---

class: center, middle

# Thank you!

Slides can be viewed at https://ecohealthalliance.github.io/odk_quality_control or PDF version downloaded at https://ecohealthalliance.github.io/odk_quality_control/odk_quality_control.pdf

R scripts for slides available at https://github.com/ecohealthalliance/odk_quality_control

R script demonstrating how to access data from **ODK Central** using `{ruODK}` package at https://github.com/ecohealthalliance/odk_quality_control/blob/main/odk_quality_control.R
