---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Slide Deck for the Presentation on Open Data Kit Quality Control Workflows for the Methods and Modelling Meetings (M3)

<!-- badges: start -->
<!-- badges: end -->

This repository contains the `Rmarkdown` script and related materials used to produce the slide deck for the presentation on Open Data Kit Quality Control Workflows for EcoHealth Alliance's **Methods and Modelling Meetings (M3)** on the 16th of November 2021.

## The slide deck

The slide deck was created using [Yihui Xie's `{xaringan}` package](https://github.com/yihui/xaringan) and Garrick Aden-Buie's ninja-themed presentation Rmarkdown template from his [`{xaringanthemer}` package](https://github.com/gadenbuie/xaringanthemer).

The **Rmarkdown** document is named `opendatakit.Rmd`. The **Rmarkdown** document relies on additional resources to produce the slide deck. These are:

* `xaringan-themer.css` - this is the CSS that comes included when using the [`{xaringanthemer}` package](https://github.com/gadenbuie/xaringanthemer)'s ninja-themed presentation template. This CSS file is dynamically re-generated using new style specifications used in the **Rmarkdown** document

* `libs` folder - this directory contains javascript libraries used by [`{xaringanthemer}` package](https://github.com/gadenbuie/xaringanthemer) to generate the HTML slides

* `images` folder - contains graphics used in the slides

These abovementioned four files are what is needed to reproduce the slide deck.

## Publishing the slide deck online

To publish the slide deck online, the following workflow using [GitHub Pages](https://pages.github.com) was used:

1. Rendered the `odk_quality_control.Rmd` file into an HTML file with the filename `index.html`. 

```{r render, echo = TRUE, eval = FALSE}
rmarkdown::render(input = "odk_quality_control.Rmd", output_file = "index.html")
```

By using `index.html` as the output filename, the resulting URL will need to just point to the enclosing directory.

2. Created a `gh-pages` branch for the `odk_quality_control` repository.

3. Enabled [GitHub Pages](https://pages.github.com) by selecting the `gh-pages` branch as the [GitHub Pages](https://pages.github.com) source.

After this, the slide deck is now published online at https://ecohealthalliance.github.io/odk_quality_control/

## Creating a PDF version of the slide deck

The slide deck was converted into PDF using the `{pagedown}` package:

```{r chrome-print, echo = TRUE, eval = FALSE}
pagedown::chrome_print(input = "index.hml", output = "odk_quality_control.pdf")
```

This operation produces a PDF file of the presentation called `odk_quality_control.pdf`. This operation requires that Google Chrome or Chromium (for Linux) is installed on your computer.

## Additional resources

In addition to the four files or directories discussed above, a file named `odk_quality_control.R` is also included in this repository. This is an R script detailing the example data quality control workflow used in this presentation.

